Measurement and Probability in Fuzzy Quantum Theories

### 2016-11-29 proposal/prop.tex
final...

\paragraph*{Research Question.} A long term, ultimate, goal of this
research proposal would be to combine the various pieces into a
coherent computational model of quantum mechanics that keeps explicit
track of resources needed to represent the interactive system
representing the quantum state, its evolution, the observables, and
the probabilities. Such a computational model of quantum mechanics
would encompass classical (reversible) computation and potentially
revolutionize quantum thinking with finite resources. The explicit
manipulation of resources enables one of key idea that emerged in our
previous work: the resources used to represent the quantum system are
in principle unrelated to the resources available to the various
observers. We conjecture that this asymmetry may prove key to
resolving some of the quantum paradoxes.

The nature of quantum mechanics realized in terms of discrete fields
and finitely-communicable probability estimates is amenable to direct
computational exploration for reasonable field sizes and observational
accuracy. We have established a set of utility libraries in
Mathematica that supports a wide variety of such numerical
experiments. Concurrently to this effort, we are also building on our
previous
work~\cite{VDS2009,AGVS2007,AGVS2005,VAS2006,finiteQC,Sabry:2003:MQC:871895.871900}
to build semantically well-founded implementations of the
computational models investigated in this proposal.  The aim of these
implementations is two-fold: (i) to provide an experimental testbed to
simulate physical phenomena (e.g.~\cite{OrtizEtAl2001,SommEtAl2002})
that complements the Mathematica libraries in which quantum algorithms
can be simulated; and (ii) to establish connections between quantum
programming concepts and classical programming concepts.  We plan to
make our software available for educational and research purposes.

>>>>

A long term, ultimate, goal of this research proposal would be to
combine the various pieces into a coherent computational model of
quantum mechanics that keeps explicit track of resources needed to
represent the interactive system representing the quantum state, its
evolution, the observables, and the probabilities. Such a
computational model of quantum mechanics would encompass classical
(reversible) computation and potentially revolutionize quantum
thinking with finite resources. The explicit manipulation of resources
enables one of key idea that emerged in our previous work: the
resources used to represent the quantum system are in principle
unrelated to the resources available to the various observers. We
conjecture that this asymmetry may prove key to resolving some of the
quantum paradoxes.  The nature of quantum mechanics realized in terms
of discrete fields and finitely-communicable probability estimates is
amenable to direct computational exploration for reasonable field
sizes and observational accuracy. We have established a set of utility
libraries in Mathematica that supports a wide variety of such
numerical experiments. Concurrently to this effort, we are also
building on our previous work~\cite
{VDS2009,AGVS2007,AGVS2005,VAS2006,finiteQC,Sabry:2003:MQC:871895.871900}
to build semantically well-founded implementations of the
computational models investigated in this proposal.

\paragraph*{Research Questions.} The aim of these implementations
include:
\begin{itemize}
\item Provide an experimental testbed to simulate physical phenomena
  (e.g.~\cite {OrtizEtAl2001,SommEtAl2002}) that complements the
  Mathematica libraries in which quantum algorithms can be simulated;
\item Establish connections between quantum programming concepts and
  classical programming concepts. We plan to make our software
  available for educational and research purposes; and 
\item Analyze the computational complexity of quantum theories and
  simulations based on discrete fields. What other (apart from
  UNIQUE-SAT) supernatural algorithms can be implemented in
  computational models based on discrete fields?
\end{itemize}
